### 2021-03-26 chapters/sections/14-technological_integration.tex
finished

This section describes how the messaging system described as part of the previous chapter can be reused for the IMP solution described in this chapter.

@@

Focus of the technological integration is to use the messaging system as described in section 4.2 to integrate the new IMP solution into the different system architecture of institutions. The layout of the messaging system should not change and the purpose of each module should stay the same. In order to adapt the architecture to the new system architecture and IMP solution, the modules can be configured and named differently.

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.6]{Diagrams/Integration Architecture 2/Technological Integration/1. Integration Overview.pdf}
    \caption{Integration Overview}
    \label{integration2:integration_overview}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.6]{Diagrams/Integration Architecture 2/Technological Integration/3. Application Overview.pdf}
    \caption{Service Application Overview}
    \label{integration2:application_overview}
\end{figure}

Institutions are able to process a predetermined set of administrative services. For each service, the institution can create a relationship template once and store it as QR code on their web page.

Users visiting an administration portal can select an administrative service, specify their home address and be forwarded to the web page of the respective institution. After scanning the appropriate QR code with the IMP client, the user can fill in the requested attributes and send the relationship request. After the relationship is established, the institution is notified and starts processing the application utilizing the shared attributes.

@@

The layout of the messaging system is identical to the one presented in section 4.2. Exactly the same type of messages are exchanged with however a slightly different purpose and content. Therefore, in this and the following sections, only important differences between the operation of the messaging system will be described.

On the "Relationship to Identity Integration Layer", modules exist for establishing the different types of application relationships. For simplification, modules only for an exemplary application process "Application A" are shown. Although the content and purpose of relationships in this case are different to the relationships of chapter 4, they are processed exactly the same: each type of relationship request is processed by an individual application request module and each type of established relationship is processed by an individual application relationship module. As the purpose of relationships in this case are applications for administrative services, each module has to be configured appropriately. For an established relationship, instead of issuing the existing system architecture to create a user profile, the application relationship module issues the existing system architecture to process an application based on the attributes shared as part of the relationship. And instead of a user profile ID, the existing system architecture responds with an application process ID. This application process ID is used in the same way as the user profile ID in section 4.2. Through the "New Relationship Mapping" channel, it is sent to the relationship mapping module, which maps each relationship to an application process and vice versa.

On the "Communication Integration Layer", the same types of messages are exchanged. However instead of user profile IDs, relationship IDs are mapped to application process IDs. In chapter 4, each module connected to the messaging adapter of the institution system was designed to be able to distinguish between different types of user profile IDs in order to for example distinguish an enterprise profile from the profile of a private person. In this case, modules can distinguish between different application types based on the application process ID.

@@

@@

Relationship templates are created manually and stored as a QR-code image on the web page of the institution. For each application type, a separate application template module can be included which stores the configuration for the template type and issues the connector to create new relationship templates. Instead of the web page, the institution system is connected, as application templates are not dynamically created for user requests but once by an employee of the institution. Each time the content of an application template has to be changed, the institution creates a new one through this module.

@@

For each type of application relationship, a different application request module is used to separate the processing of requests for different application types into application request processing modules. Depending on the type of application, different requirements regarding the acceptance of application requests exist. A request processing module might for example check if the home address of the user is within the are of responsibility of the institution. Only after the relationship and therefore the application request is accepted the application will be processed.

@@

For each type of application relationship a different application relationship module is used for processing established relationships. The institution system is notified of an accepted application through the "Process Application Request" channel. As part of the existing system architecture, institution systems received applications through the data exchange platform as an application document. In order to adhere to this standard, the message translator connected to the "Process Application Request" channel translates the relationship content into the application format of documents delivered through the data exchange platform. As explained in section 2.3.2, form servers create the application documents based on a standard defined by FIM.

Institution systems are assumed to uniquely reference active applications with process IDs. After starting the application process, the institution responds through the messaging adapter with the process ID on the "Application Process Reference Response" channel. This process ID is then published on the "New Relationship Mapping" channel. The result is, that in contrast to the messaging system of chapter 4, relationships are not mapped to user profiles but to application processes.

@@

As an example for a message type processed on the "Communication Integration Layer", the application attribute change messages are selected.

In the same way as in chapter 4, attribute change message types are published on individual channels, this time however, instead of a user profile ID, they contain a process ID. As described on the example of different user profile ID types in chapter 4, different process ID types may exist in this case too. For example, an attribute change request for an application of type "A" might be processed differently to an attribute change request for an application of type "B". To process different types of attribute change requests separately, multiple application attribute change request modules can be included, each of them filtering for different types of process IDs. If it is possible to retrieve the application type from the process ID, processing of each application type could be separated into multiple application change request modules. In case application types are not retrievable from process IDs, an additional module can be included, which requests the application type from the institution system based on the process ID. An alternative would be for the application relationship module to not only publish the process ID corresponding to a relationship ID on the "New Relationship Mapping" channel but also the application type.

Only after the attribute change request is accepted, the institution system will be notified about the attribute change.

@@

Continuing the example of how application attribute change messages are processed on the "Communication Integration Layer", the application attribute sync module processes messages on the "Attribute Change Established" channel. With the same reasoning as in the previous section, the message filter attached to the channel enables to include multiple application attribute sync modules, processing different application types.

@@

Similar to the attribute change messages, mail messages can be fitted to the process based scenario, enabling institutions to send mails to a process ID which the messaging system translates to a relationship ID in order for the correct IMP identity to receive it. In addition to the one directional message transfer from the institution to the IMP identity, the application mail module and the mail message module could be expanded to be able to receive mails from an IMP identity through an established relationship. The relationship ID of the message would be mapped to the corresponding process ID and submitted to the institution system. The institution system could then send the mail to the responsible employee.

@@

Authentication messages and modules are not required for this IMP solution, as no user profile and therefore no login process exists. However, depending on the use case, authentication messages could be included as an additional accept / reject communication method between institution and user.

\subsection{Evaluation}
The integration architecture enables users to directly interact with the system architecture of the Service Provider for submitting applications. This eliminates the need for administration portals to process personal information.

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.6]{Diagrams/Integration Architecture 2/IMP Personal Information.pdf}
    \caption{IMP System Architecture Personal Information}
    \label{integration2:imp_personal_information}
\end{figure}

Comparing the existing system architecture of the OZG with the system architecture utilizing IMP, the amount of systems processing personal information is reduced. Using this IMP solution, only the user and the responsible institution have access to personal information.

The integration architecture described in section 4.2 was able to integrate into two different system architectures. This demonstrates, that the presented integration architecture is applicable to a variety of system architectures.

It was also able to integrate a user oriented (persistent) relationship scenario as well as a process oriented (temporary) relationship scenario. This demonstrates that the presented integration architecture can be used for integrating a variety of IMP solutions.

The modular message based approach made it possible to use the integration architecture in a different scenarios, by enabling easy configuration of modules for new requirements.

>>>>

This section presents how the messaging system described as part of the previous chapter can be reused for the IMP solution of this chapter.

@@

The focus of the technological integration is to use the messaging system as described in section 4.2 to integrate the new IMP solution into the different system architecture of institutions. The layout of the messaging system should not change and the purpose of each module should stay the same. To adapt the architecture to the new system architecture and IMP solution, the modules can be configured and named differently.

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.6]{Diagrams/Integration Architecture 2/Technological Integration/1. Integration Overview.pdf}
    \caption{Integration Overview}
    \label{integration2:integration_overview}
\end{figure}

@@

@@

As shown in figure \ref{integration2:messaging_overview}, the layout of the messaging system is identical to the one presented in section 4.2. The same type of messages are exchanged with a slightly different purpose and content. Therefore, in this and the following sections, only the important differences between the operation of the messaging system will be detailed.

On the "Relationship to Identity Integration Layer" modules exist for establishing different types of application relationships. For simplification, only modules for an exemplary application process "Application A" are shown. Although the content and purpose of the relationships in this case are different to the relationships of chapter 4, they are processed the same: each type of relationship request is processed by an individual application request module and each type of established relationship is processed by an individual application relationship module. As the purpose of relationships in this case are applications for administrative services, each module has to be configured appropriately. For an established relationship, instead of issuing the existing system architecture to create a user profile, the application relationship module issues the existing system architecture to process an application based on the attributes shared as part of the relationship. Moreover, instead of a user profile ID, the existing system architecture responds with an application process ID. This application process ID is used in the same way the user profile ID is utilized in section 4.2. Through the "New Relationship Mapping" channel, it is sent to the relationship mapping module, which maps each relationship to an application process and vice versa.

On the "Communication Integration Layer" the same types of messages are exchanged. However, instead of user profile IDs, relationship IDs are mapped to application process IDs. In chapter 4, each module connected to the messaging adapter of the institution system was designed to be able to distinguish between different types of user profile IDs to, for example, distinguish an enterprise profile from the profile of a private person. \\
In this case, modules can distinguish between different application types based on the application process ID.

\subsection{Application Process}

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.6]{Diagrams/Integration Architecture 2/Technological Integration/3. Application Overview.pdf}
    \caption{Administrative Service Application Overview}
    \label{integration2:application_overview}
\end{figure}
This section describes the process of applying for an administrative service as shown in figure \ref{integration2:application_overview}.
Institutions are able to process a predetermined set of administrative services. For each service, the institution can create a relationship template once and store it as QR code on their web page. 
Users visiting an administration portal can select an administrative service, specify their home address, and be forwarded to the web page of the respective institution. After scanning the appropriate QR code with the IMP client, the user can fill in the requested attributes and send the relationship request. After the relationship is established, the institution is notified and starts processing the application utilizing the shared attributes.

@@

In this section, each messaging module mentioned in the overview of the message architecture is described.

@@

Relationship templates are created manually and stored as a QR-code on the web page of the institution. For each application type, a separate application template module can be included, which stores the configuration for the template type and issues the connector to create new relationship templates. Instead of the web page, the institution system itself is connected, as application templates are not dynamically created for user requests but only once by an employee of the institution. Each time the content of an application template has to be changed, the institution creates a new one with this module.

@@

For each type of application relationship, a different application request module is used to separate the processing of requests for different application types into application request processing modules. Depending on the type of application, different requirements regarding the acceptance of application requests exist. A request processing module might, for example, check if the home address of the user is within the area of responsibility of the institution. Only after the relationship and therefore the application request is accepted, the application will be processed.

@@

For each type of application relationship, a different application relationship module is used for processing the established relationships. The institution system is notified of an accepted application through the "Process Application Request" channel. As part of the existing system architecture, institution systems receive applications through the data exchange platform as an application document. To adhere to this standard, the message translator connected to the "Process Application Request" channel translates the relationship content into the application format of documents delivered through the data exchange platform.

Institution systems are assumed to uniquely reference active applications with process IDs. After starting the application process, the institution responds through the messaging adapter with the process ID on the "Application Process Reference Response" channel. This process ID is then published to the "New Relationship Mapping" channel. In contrast to the messaging system of chapter 4, the relationships in this case are not mapped to user profiles but to application processes.

@@

As an example, for a message type processed on the "Communication Integration Layer", the application attribute change messages are selected.

Similar to the process in chapter 4, attribute change message types are published on individual channels, this time however, instead of a user profile ID, they contain a process ID. As described in the example of different user profile ID types in chapter 4, different process ID types may exist in this case too. For example, an attribute change request for an application of type "A" might be processed differently to an attribute change request for an application of type "B". To process different types of attribute change requests separately, multiple application attribute change request modules can be included, each of them filtering for different types of process IDs. If it is possible to retrieve the application type from the process ID, the processing of each application type could be separated into multiple application change request modules. For the case that application types are not retrievable from process IDs, an additional module can be included, which requests the application type from the institution system based on the process ID. An alternative way would be for the application relationship module to not only publish the process ID corresponding to a relationship ID on the "New Relationship Mapping" channel but also the application type.

Only when the attribute change request is accepted, the institution system will be notified about the attribute change.

@@

Continuing the example of how application attribute change messages are processed on the "Communication Integration Layer", the application attribute sync module processes messages on the "Attribute Change Established" channel. With the same reasoning as in the previous section, the message filter attached to the channel enables to include multiple application attribute sync modules, each processing different application types.

@@

Similar to attribute change messages, mail messages can be fitted to the process-based scenario, enabling institutions to send mails to a process ID which the messaging system translates to a relationship ID in order for the correct IMP identity to receive it. In addition to the one-directional message transfer from the institution to the IMP identity, the application mail module and the mail message module could be expanded to be able to receive mails from an IMP identity through an established relationship. The relationship ID of the message would be mapped to the corresponding process ID and submitted to the institution system. The institution system could then send the mail to the responsible employee.

@@

Authentication messages and modules are not required for this IMP solution, as no user profile and therefore no login process exists. However, depending on the use case, authentication messages could be included as an additional accept / reject communication method between the institution and user.

\subsection{Evaluation}

@@

The integration architecture enables users to directly interact with the system architecture of the Service Provider for submitting applications. This eliminates the need for administration portals to process personal information. As shown in figure \ref{integration2:ozg_personal_information}, every component and every domain of the existing OZG system architecture processes personal information. Using the IMP solution presented in this chapter, only the user and the responsible institution have access to personal information, which improves data protection.

The integration architecture described in section 4.2 was able to integrate into two different system architectures. This demonstrates that the presented integration architecture is applicable to a variety of system architectures. In addition, it was able to integrate a user-oriented (persistent) relationship scenario as well as a process-oriented (temporary) relationship scenario. This demonstrates that the presented integration architecture can be used for integrating a variety of IMP solutions. The modular message-based approach makes it possible to use the integration architecture in different scenarios by enabling easy configuration of modules for new requirements.

### 2021-03-26 chapters/sections/15-conclusion.tex
Einleitung überarbeitet
finished

Finally, this chapter aims to summarize the most important results of the thesis and give an outlook on possible future developments and contributions.

@@

\begin{itemize}
    \item Purpose of the thesis is to design integration solutions for identity management provisioning systems on the example of OZG.
    \item The integration is separated into a business scenario integration trough an IMP solution and a system integration trough a messaging system.
    \item Trough two IMP solutions, it was shown, how IMP can be utilized in different identity management scenarios
    \item In the first scenario, an IMP solution was presented which focused on enhancing usability of 
    \item The thesis introduced identity management provisioning as an identity management service capable of replacing conventional user profiles
    \item On the example of the OZG, the thesis showed 
\end{itemize}

@@

\begin{itemize}
    \item experimental evaluation
    \item 
\end{itemize}

>>>>

Finally, this chapter summarizes the most important results of this thesis and gives an outlook on possible future developments and contributions.

@@

The bachelor thesis uses the scenario of the OZG as a small-scale model for the general problem of distributed partial identities. Focus is the utilization and system integration of IMP for Service Providers with existing user profile-based system architectures.

The first objective of the bachelor thesis is to examine the problems of user profile-based identity management and interoperable OZG user profiles and to describe to what extent IMP can be a solution. In chapter \ref{chapter:user_profile_identity_management} the thesis describes multiple problems of user profile-based identity management regarding usability, data protection and security. They lack usability as they lead to fragmentation into partial identities, accessible only through separate applications and web pages. Data protection is reduced due to the lack of oversight for the user about existing user profiles and shared personal information. The distribution of personal information to various Service Providers increases the risk of data being stolen. Especially in the case of the OZG, issues regarding data protection exist: \\
As a result of the "one-for-all" method, access to administrative services is distributed across multiple administration portals. Users applying for services therefore have to share large quantities of personal information with the administration portals of multiple federal states. As it is the institutions and not the administration portals that eventually process the personal information, sharing personal information directly with institutions would be favourable regarding data minimisation. \\
Possibilities for solving the mentioned problems through IMP are presented. IMP enables to increase usability by enabling users to interact, manage, and share personal information with Service Providers through only one application. Data protection is increased, as users can keep an overview of shared personal information. Security is improved, as IMP provides secure communication channels. In the case of the OZG, IMP is a possibility to increase usability and even eliminate the need for administration portals to process personal information, favouring data minimization. Based on the mentioned problems and solutions, the chapter concludes that IMP can be taken into consideration for solving problems of fragmented partial identities.

The bachelor thesis therefore moves to the second objective to design "IMP solutions" which describe the possibilities of how IMP services and tools can be specifically utilized on the example of - but not limited to - the OZG use case. The thesis presents two solutions for IMP utilization for Service Providers with existing user profile-based identity management. The first solution utilizes IMP in addition to existing user profiles. It provides an optional way for users to create and manage OZG user profiles through an IMP client. \\
This is done by enabling users to create new user profiles through existing IMP identities by establishing an IMP relationship. By exchanging IMP messages, users are able to access and manage their user profile. The IMP solution especially improves usability as OZG user profiles can be managed along with many other user profiles of other Service Providers through the same IMP client. The solution requires integration only with the administration portal component and can operate without disrupting the existing operation of the system architecture.

The approach of enhancing accessibility to existing user profiles through an IMP client can be used to replace the need for OZG user profiles to be interoperable. Users can effortlessly create, manage, and login to user profiles for each federal state and administration portal. In addition to that, not only user profiles in the context of the OZG can be managed through IMP but user profiles of any integrated Service Provider. This IMP utilization therefore is a possibility to solve the usability problems of partial identities.

However, this IMP solution does not solve many of the data protection issues, as the user profile-based identity management system remains in operation. Therefore, the bachelor thesis presents a second IMP solution, with a focus on data protection.

The second solution utilizes IMP to replace user profiles. Instead of persistently sharing personal information with administration portals in the form of user profiles, IMP is used to temporarily share personal information with institutions for the period of time where an application is being processed. This is done by enabling users to establish different types of IMP relationships, each corresponding to a different application type. This IMP solution especially improves data protection, as only the institution responsible for processing the application receives personal information. In addition to that, sharing personal information only for a short period of time decreases the risk of personal information being misused or stolen.

The approach of replacing user profiles with temporary sharing of personal information for individual business processes can be used to eliminate partial identities while increasing data protection. However, while this approach might be usable in the context of the OZG, Service Providers in the private sector might not be interested in switching to this fundamentally different approach of identity management. Not as a result of technological problems but due to corporate policy reasons ("Never change a running system"). The advantage of IMP is: users can access Service Providers utilizing either IMP solutions through their IMP application.

The third objective of the bachelor thesis is to design a technological integration architecture for the integration of IMP solutions into existing system architectures of Service Providers. The thesis presents a modular messaging system, which enables the integration of IMP solutions into existing system architectures of Service Providers. Through the messaging system, the system architecture of the Service Provider can establish multiple types of IMP relationships and define, exchange, and process different message types. The messaging system maps relationships to entities that the SP systems understand. This can be, for example, the mapping of an IMP relationship to a user profile or to an application process. It also translates between the data models of IMP and Service Provider. As a result of using messaging and messaging modules, the integration system can be configured and extended to fit the requirements of different IMP solutions and system architectures of Service Providers.

The messaging system is shown, integrating both previously described IMP solutions into different system architectures. The first IMP solution was integrated into the administration portal component, the second IMP solution was integrated into the institution domain. This demonstrates that the messaging system is capable of integrating different IMP solutions into different system architectures.

@@

The bachelor thesis describes the integration of IMP on a theoretical level. Developing a prototype of the messaging system as part of an experimental evaluation can give further information about its feasibility, performance, and stability. The prototype can furthermore be used to test integration of the first IMP solution into the OZG system architecture of a federal state, to validate the functionality of the IMP solution and the messaging system, and to find missing features. Due to the modular design of the messaging system, new messaging modules implementing missing features can be added.

To make the modular approach of the integration architecture more utilizable for Service Providers, a "Module Store" can be hosted, where enterprises can publish messaging modules for integration of enterprise systems they sell. Through an IMP tool, Service Providers could be enabled to generate a custom messaging system with messaging modules for selected systems. With "only" configuration of the messaging modules remaining, this could reduce the time required for implementing and deploying the integration architecture.

The thesis presents OZG as a Service Provider relevant for the integration of IMP.
Identity Management Provisioning, however, is relevant for many other Service Providers. For example, in the scope of the National Education Platform, "a digital portal is being created that in the future will support universities as well as national and international students who are seeking a stay abroad, a course of study, or a change of study location in Germany along their educational career through a digital offering that is interoperable across institutions" \cite{digitale_bildung}. In accordance with IMP solutions in the thesis, an IMP solution for this new use case can be designed and the presented messaging system can be configured and expanded to suit new requirements and to be used for system integration.

### 2021-03-26 chapters/sections/2-objective.tex
finished

Purpose of the thesis is to design an integration solution for identity management provisioning and identity management provisioning systems.

In order to make the integration suitable to real life requirements, the OZG is selected as an example due to its current relevance and high requirements regarding identity management.

Two solutions for IMP integration and one message based IMPS integration architecture are presented.

The IMP solutions describe two important fundamentally different approaches of how service providers can utilize IMP for identity management. In the first solution IMP is used alongside existing user profiles, utilizing IMP as a way of improving existing user profile based identity management solutions without replacing them. This approach is especially important for service providers in the private sector which require integration with minimal risk and cost.

In the second solution, IMP is used as an alternative way for identity management without user profiles. This approach is especially important for service providers constructing new system architectures or have strong interest in data protection and data economy.

The message based IMPS integration architecture is designed to enable integration of multiple IMP solutions into different system architectures of service providers. This is demonstrated by presenting the integration of both IMP solutions contained in the thesis, each into a different system architecture.

>>>>

The bachelor thesis uses the scenario of the OZG as a small-scale model for the general problem of distributed partial identities. Focus is the utilization and system integration of IMP for Service Providers with existing user profile-based system architectures. A federal state, in the context of the OZG, is used as an example for a Service Provider.

One objective of the bachelor thesis is to examine the problems of user profile-based identity management and interoperable OZG user profiles and to describe to what extend IMP as an approach can be a solution.

Another objective of the bachelor thesis is to design "IMP solutions" which describe the possibilities of how IMP services and tools can be concretely utilized on the example of - but not limited to - the OZG use case. IMP solutions demonstrate how IMP relationships and IMP messages can be utilized by users and Service Providers to share personal information and interact to solve the problems of distributed partial identities related to usability, data protection, and security.

Using OZG as an example, another objective of the bachelor thesis is to design a technological integration architecture for the integration of IMP solutions into existing system architectures of Service Providers. Its integration capabilities, however, can not be limited to one IMP solution and one system architecture. On the contrary, it has to be capable of integrating various IMP solutions into different system architectures. The integration system therefore has to be configurable and expandable according to the requirements of new IMP solutions and system architectures. To demonstrate the capabilities of the integration architecture in theory, IMP solutions presented as part of the thesis have to be integrated into different system architectures in the context of the OZG.

IMP solutions and integration architecture presented in the bachelor thesis should not be perfectly fitted to the scenario of the OZG but have to also be utilizable by Service Providers in other contexts.

### 2021-03-26 chapters/sections/3-structure_of_work.tex
finished

The chapter "Background and Related Work" covers important topics that following chapters rely on. 
The chapter "User Profile Identity Management" describes disadvantages of user profile based identity management in general as well as in context of the OZG. Possible solutions through IMP are described. 
The chapter "IMP Utilization as Extension" presents the first IMP solution as well as its integration through the technological integration architecture.
The chapter "IMP Utilization as Replacement" presents the second IMP solution as well as its integration through the technological integration architecture.
The chapter "Conclusion" summarizes the most important results of the thesis and mentions possible future developments and contributions.

>>>>

The chapter "Background and Related Work" covers important topics that the following chapters rely on. 
The chapter "User Profile Identity Management" describes the disadvantages of user profile-based identity management in general as well as in the context of the OZG. Possible solutions through IMP are described. 
The chapter "IMP Utilization as Extension" presents the first IMP solution as well as its integration through the technological integration architecture.
The chapter "IMP Utilization as Replacement" presents the second IMP solution as well as its integration through the technological integration architecture.
The chapter "Conclusion" summarizes the most important results of the thesis and mentions possible future developments and contributions.
